---
output: github_document
---

```{r echo=FALSE}
options(digits = 2)
knitr::opts_chunk$set(message = FALSE, fig.path = 'inst/README_files/')
set.seed(1234)
```

# <img src='man/figures/logo.svg' align="right" height="139" /> R package recommenderlab - Lab for Developing and Testing Recommender Algorithms 


```{r echo=FALSE, results = 'asis'}
pkg <- 'recommenderlab'

library(stringr)

  cat(str_interp("[![CRAN version](http://www.r-pkg.org/badges/version/${pkg})](https://CRAN.R-project.org/package=${pkg})\n"))
  cat(str_interp("[![stream r-universe status](https://mhahsler.r-universe.dev/badges/${pkg})](https://mhahsler.r-universe.dev/ui#package:${pkg})\n"))
  cat(str_interp("[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/grand-total/${pkg})](https://CRAN.R-project.org/package=${pkg})\n"))
```

Provides a research infrastructure to develop and evaluate collaborative filtering recommender algorithms. This includes a sparse representation for user-item matrices, many popular algorithms, top-N recommendations, and cross-validation. 
The package supports rating (e.g., 1-5 stars) and unary (0-1) data sets. Supported algorithms are:
    
* User-based collborative filtering (UBCF)
* Item-based collborative filtering (IBCF)
* SVD with column-mean imputation (SVD)
* Funk SVD (SVDF)
* Alternating Least Squares (ALS)
* Matrix factorization with LIBMF (LIBMF)
* Association rule-based recommender (AR)
* Popular items (POPULAR)
* Randomly chosen items for comparison (RANDOM)
* Re-recommend liked items (RERECOMMEND)
* Hybrid recommendations (HybridRecommender)

For evaluation, the framework supports given-n and all-but-x protocols with

* Train/test split
* Cross-validation
* Repeated bootstrap sampling

Evaluation measures are:

* Rating errors: MSE, RMSE, MAE
* Top-N recommendations: TPR/FPR (ROC), precision and recall

## Installation

__Stable CRAN version:__ install from within R with
```{r eval=FALSE}
install.packages("recommenderlab")
```

__Current development version:__ Install from [r-universe.](https://mhahsler.r-universe.dev/ui#package:recommenderlab)

## Usage

Load the package and prepare a dataset (included in the package). The MovieLense
data contains user ratings for movies on a 1 to 5 star scale.
We only use here users with more than 100 ratings.
```{r}
library("recommenderlab")
data("MovieLense")

MovieLense100 <- MovieLense[rowCounts(MovieLense) > 100, ]
MovieLense100
```

Train a user-based collaborative filtering recommender using a small training set.
```{r}
train <- MovieLense100[1:100]
rec <- Recommender(train, method = "UBCF")
rec
```

Create top-N recommendations for new users (users 101 and 102).
```{r}
pre <- predict(rec, MovieLense100[101:102], n = 5)
pre
```

```{r}
as(pre, "list")
```

Use a 10-fold cross-validation scheme to compare the top-N lists of several algorithms.
Movies with 4 or more stars are considered a good recommendation. 
We plot true negative vs. true positive rate for top-N lists of different lengths.
```{r TNR_vs_TPR}
scheme <- evaluationScheme(MovieLense100, method = "cross-validation", k = 10, 
  given = -5, goodRating = 4)
scheme

algorithms <- list(
  "random items" = list(name = "RANDOM", param = NULL),
  "popular items" = list(name = "POPULAR", param = NULL),
  "user-based CF" = list(name = "UBCF", param = list(nn = 3))
)

results <- evaluate(scheme, algorithms, type = "topNList",
  n=c(1, 3, 5, 10), progress = FALSE)

plot(results, annotate = 2, legend = "topleft")
```

## Shiny App

A simple Shiny App running recommenderlab can be found at  [https://mhahsler-apps.shinyapps.io/Jester/](https://mhahsler-apps.shinyapps.io/Jester/) 
([source code](https://github.com/mhahsler/recommenderlab/tree/master/Work/apps)). 


## References

* Michael Hahsler (2022) recommenderlab: An R framework for developing and testing recommendation algorithms. arXiv:2205.12371 [cs.IR]. DOI: [10.48550/arXiv.2205.12371](https://doi.org/10.48550/arXiv.2205.12371).
* recommenderlab [reference manual](https://CRAN.R-project.org/package=recommenderlab/recommenderlab.pdf)
*  Suresh K. Gorakala and Michele Usuelli (2015) [Building a Recommendation System with R](https://www.amazon.com/Building-Recommendation-System-Suresh-Gorakala/dp/1783554495) (Packt Publishing) features the package recommenderlab.
