---
title: "Getting started with recommenderlab"
author: "Michael Hahsler"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with recommenderlab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(recommenderlab)
set.seed(1234)
```

`recommenderlab` provides tools for representing user–item data, fitting
recommendation algorithms, producing recommendations, and evaluating their
quality. This vignette walks through that workflow using the package's bundled
MovieLense ratings data.

## Installation

Install the released package from CRAN, then load it in your R session:

```{r install, eval=FALSE}
install.packages("recommenderlab")
```

```{r load-package}
library(recommenderlab)
```

## Load and prepare ratings

The `MovieLense` data contains ratings on a one-to-five-star scale. It is stored
as a sparse `realRatingMatrix`: users are rows, movies are columns, and missing
ratings are not stored as zeros. We select users who rated more than 100 movies
to give the recommendation algorithms enough information to work with.

```{r data}
data("MovieLense")
MovieLense

MovieLense100 <- MovieLense[rowCounts(MovieLense) > 100, ]
MovieLense100
```

Basic summaries help describe the data before modeling. For example,
`rowCounts()` counts ratings per user, and `getRatings()` extracts the observed
rating values.

```{r inspect-data}
summary(rowCounts(MovieLense100))
summary(getRatings(MovieLense100))
```

## Fit a recommender and make recommendations

`Recommender()` learns a model from a training rating matrix. Here we fit
user-based collaborative filtering (UBCF) on the first 300 selected users.
`predict()` then produces a top-five list for two other users. The default
output is a `topNList`; coerce it to a list to see the recommended movie
titles.

```{r recommendations}
train <- MovieLense100[1:300, ]
rec <- Recommender(train, method = "UBCF")
rec

recommendations <- predict(rec, MovieLense100[301:302, ], n = 5)
recommendations
as(recommendations, "list")
```

The package also supports predicted ratings. Request `type = "ratings"` when
the numeric estimates are more useful than a ranked list.

```{r predicted-ratings}
predicted_ratings <- predict(
  rec,
  MovieLense100[301:302, ],
  type = "ratings"
)
as(predicted_ratings, "matrix")[, 1:6]
```

## Evaluate recommendations

Evaluation should simulate the information available when recommendations are
made. An all-but-five scheme withholds five ratings per user and uses the
remaining ratings as known input. Here, ratings of four stars or higher count
as positive feedback. `evaluationScheme()` supports train/test splits,
cross-validation, and bootstrap evaluation.

```{r evaluation-scheme}
evaluation_data <- MovieLense100[1:200, ]
scheme <- evaluationScheme(
  evaluation_data,
  method = "cross-validation",
  k = 5,
  given = -5,
  goodRating = 4
)
scheme
```

Compare a popularity-based recommender with a random baseline. `evaluate()`
fits each method on every training fold, creates top-N recommendations, and
calculates measures from the withheld ratings. The resulting true-positive and
false-positive rates can be plotted to compare recommendation list lengths.

```{r evaluate}
algorithms <- list(
  `popular items` = list(name = "POPULAR", param = NULL),
  `random items` = list(name = "RANDOM", param = NULL)
)

results <- evaluate(
  scheme,
  algorithms,
  type = "topNList",
  n = c(1, 3, 5, 10),
  progress = FALSE
)
getResults(results[[1]])
```

Plot the average true-positive rate against the false-positive rate for each
recommendation list length:

```{r plot-results, fig.width=7, fig.height=5}
plot(results, annotate = TRUE, legend = "topleft")
```

For predicted ratings, `evaluate()` can instead report rating error measures
such as RMSE, MSE, and MAE by using `type = "ratings"`. See
`?calcPredictionAccuracy` for the measures available for direct predictions
and `?evaluate` for details on evaluation results.

## Where to go next

The package includes additional algorithms such as item-based collaborative
filtering (IBCF), matrix factorization, association-rule recommenders, and
hybrid recommenders. Use `recommenderRegistry$get_entry_names()` to see the
methods available in your installation. The reference manual documents each
algorithm, data class, and evaluation helper.
